Computation and comparison of the stable Northeastern US marine boundary layer

### 2020-12-03 Paper/main.tex
Add wall model description
Added tikZ image labels
Fix typos and stuff

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{longtable,tabularx}
\usepackage{fixme}
\setlength\LTleft{0pt}
\hypersetup{
  colorlinks=true
}

@@

\author{Nathaniel B. deVelder\footnote{Postdoctoral appointee, Wind
    Energy Technologies Department, AIAA member} and Alan
  S. Hsieh.\footnote{Postdoctoral appointee, Wind Energy Technologies
    Department, AIAA member}} \affil{Sandia National Laboratories,
  Albuquerque, NM 87185}

@@

\begin{abstract}
  In this work we investigate the behavior of stable marine boundary
  layers located near the coast of the Northeastern United States.
  Using the ExaWind LES codes, three stable atmospheric conditions
  were chosen to match the Cape Wind measurements of Archer \emph{et
    al} with wind speeds of 5 m/s, 10 m/s, and 15 m/s.  The behavior
  of the stable boundary layers, including mean flow quantities and
  turbulent statistics, are examined and compared to previous
  computations of the neutral and unstable offshore boundary layer at
  the same location.  This study also examines the domain and mesh
  requirements necessary to capture the turbulent scales for the
  stable offshore boundary layers.  Finally, we compare solutions
  computed using both Nalu-Wind and AMR-Wind solvers, and report on
  their relative accuracy and performance in this study.

@@

\section{Nomenclature}

{\renewcommand\arraystretch{1.0}
\noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
$\alpha$  & Wind shear exponent  \\
$f$       & Frequency \\
$f_{max}$  & Maximum resolvable frequency due to mesh limits \\
$f_{Ny}$   & Nyquist frequency \\
$L$       & Turbulent integral length scale \\
$u_\tau$   & Friction velocity    \\
$R_{ij}$   & Two point correlation tensor \\
$TI$      & Turbulence intensity \\
$TKE$     & Turbulent kinetic energy 
\end{longtable*}}

@@

In addition to these measurement campaigns, large eddy simulations (LES) have
also been used to study ABL stability characteristics.  Recent work by
Kaul et al. \cite{kaul2020large} has shown that LES computations using
Nalu-Wind can capture the neutrally and convectively unstable onshore
ABL.  Previous simulations by Cheung et al. \cite{cheung2020large}
successfully replicated the unstable and neutral ABL corresponding to
the Cape Wind meteorological tower measurements
\cite{archer2016predominance} and showed the effects of surface
heating on atmospheric stratification. LES of the stable ABL (SBL) presents 
certain challenges compared to neutral and convective stability states, stemming 
primarily from small dominant eddy length scales and low turbulence intensity. 
These challenges require higher resolution grids and/or advanced subgrid 
scale (SGS) models. As such, LES research of the SBL has been slower 
to evolve than the other stability states.

The first successful LES of the SBL was published by Mason and Derbyshire in 
1990 \cite{Mason1990}. They used a strategy of simulating a near steady-state 
neutral stability turbulent simulation, then added a cooling-surface boundary 
to develop the stable ABL. They recorded a need for implausibly strong initial 
perturbations to sustain turbulence using the Smagorinsky SGS model. Brown et 
al.\ \cite{Brown1994} worked with Mason and Derbyshire to add stochastic backscatter 
to the SGS model, resulting in better agreement with surface layer observations. 
Andren, \cite{Andren1995}, simulated two SBL types, one with a neutral layer aloft, and 
one capped by an inversion, comparing two SGS models. He found significant improvements 
with the Sullivan SGS model, on par with the earlier backscatter model. Kemp and 
Thompson \cite{Kemp1996} used the same model as \cite{Brown1994} to study particle 
dispersion, but the LES code did not replicate an expected plume growth. Derbyshire 
wrote a review of SBL modeling along with future projections in 1999 \cite{Derbyshire1999}, 
noting the importance of local Richardson number and identifying a need to go 
to higher resolutions in SBL research.

In 2000, Kosovic and Curry \cite{Kosovic2000} used an LES simulation with a nonlinear SGS 
model containing backscatter to favorably compare to the BASE measurements, showing that 
the SBL could reach a quasi steady-state. They also showed good agreement with the Nieustadt 
analytical model. In the same year, Saiki et al. \cite{Saiki2000} simulated an SBL with 
a strong capping inversion at z=500m, producing reasonable surface layer profiles and 
identifying a need for SBL specific SGS models. Beare and MacVean \cite{Beare2004} 
presented the first grid-focused work on LES of the SBL, using grids with deltas 
between 2m and 12.5m, finding significant variation between grid sizes. The inclusion 
of a backscatter model resulted in significant variation at a 6.25m resolution, but 
very little at a 2m resolution. Basu et al. \cite{Basu2005} published a set of 
turbulence statistics, measurements, and supporting LES simulations in support 
of the viability of Nieustadt's local scaling hypothesis under very stable 
conditions. Jimenez and Cuxart \cite{Jimenez2005} set out to quantify the range 
for which standard SGS models, using the Kolmogorov theory for dissipation, were 
valid for increasingly stable flows. They found the model to perform adequately 
for weakly stable, and some moderately stable flow regimes.

An inter-comparison of LES simulations of the SBL was organized in 2006, the 
results of which are published in Beare et al.\ \cite{Beare2006}. Collectively, 
this is knows as the GABLS initiative, and involved simulating a moderately 
stable flow taken from Kosovic and Curry \cite{Kosovic2000}. Eleven codes 
were compared in all, wherein grid deltas of less than 12.5m were found to 
successfully maintain resolved turbulence, grid deltas less than 3.125m 
showed little statistical change, and results became independent of SGS 
modeling at 1m resolution. In the same year, Kumar et al. \cite{Kumar2006} 
successfully simulated an entire diurnal cycle with LES using a Lagrangian 
dynamic scale dependent SGS model. This model has been used by several 
researchers since 2011, including Lu and Porte-Agel \cite{Lu2011}, Huang 
and Bou-Zeid \cite{Huang2013}, Park et al.\ \cite{Park2014}, and 
Abkar and Porte-Agel \cite{Abkar2015}. Stoll and Porte-Agel \cite{Stoll2007} 
compared three grid refinements and three scale dependent dynamic SGS models, 
showing that the Lagrangian and locally averaged models reproduced important 
flow features. The work of van Dop and Axelsen \cite{vanDop2007} provided 
even more evidence in support Nieuwstadt's local scaling hypothesis for SBLs. 
Chow and Zhou \cite{Chow2011} showed evidence that the Smagorinsky model is 
simply too dissipative for use in strongly stable flows.

More recent work in LES of the SBL includes Matheou \cite{Matheou2016} who 
explored a large parameter space using the Smagorinsky-Lilly SGS 
closure, as well as Sullivan et al.\ \cite{sullivan2016turbulent}. Several researchers 
have applied SBL simulation in the wind energy space \cite{Aitken2014, Abkar2015, Ghaisas2017}. 
This year, Wurps et al. \cite{Wurps2020} provided a rigorous grid resolution study for the SBL, 
looking at mean flow, turbulence variables, and two-point correlations, using a dynamic 
SGS model that allows backscatter. They found grid requirements to be 2.5m, 10m, and 20m 
for stable, neutral, and convective boundary layers respectively. They advocate looking 
at two point correlations in order to quantify flow structure size, which in turn 
provides a basis for evaluating grid resolution. In all of the reviewed literature, SBL LES 
has been run in the onshore context only, a complete comparison
including offshore stable ABL conditions has yet to be completed.

\subsection{Objectives}
In the current study on stable marine boundary layers, we aim to build
on the previous work in the literature and complete several
objectives.  First, a series of stable boundary layer conditions will
be computed using the Exawind solvers which match the measurements of
Archer et al \cite{archer2016predominance}.  The behavior of the
stable boundary layers, including mean flow quantities and turbulent
statistics, are examined and compared to the neutral and unstable
cases computed in \cite{cheung2020large}.  Secondly, we examine the
domain and mesh requirements necessary to capture the turbulent scales
for the stable offshore boundary layers.  Finally, two LES solvers
within the Exawind project are compared in this study: the
unstructured, incompressible Navier-Stokes solver Nalu-Wind and the
structured, adaptive mesh refinement solver AMR-Wind.  Both codes are
used to calculate the same stable boundary layer, and the relative
differences between the codes are reported.

>>>>

\usepackage{graphicx}
\usepackage{amsmath}
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{longtable,tabularx}
\usepackage{fixme}
\setlength\LTleft{0pt}
\hypersetup{
  colorlinks=true
}
\usepackage{tikz}

\newcommand{\average}[1]{\langle #1 \rangle}

@@

\author{Nathaniel B. deVelder\footnote{Postdoctoral appointee, Wind
    Energy Technologies Department, AIAA member} and Alan
  S. Hsieh.\footnote{Senior member of technical staff, Wind Energy
    Technologies Department, AIAA member}} \affil{Sandia National
  Laboratories, Albuquerque, NM 87185}

@@

\begin{abstract}
  In this work we investigate the behavior of stable marine boundary
  layers located near the coast of the Northeastern United States.
  Using the ExaWind LES codes, three stable atmospheric conditions
  were chosen to match the Cape Wind measurements of Archer \emph{et
    al.} with wind speeds of 5 m/s, 10 m/s, and 15 m/s at the 20m
  measurement height.  The behavior of the stable boundary layers,
  including mean flow quantities and turbulent statistics, are
  examined and compared to previous computations of the neutral and
  unstable offshore boundary layer at the same location.  This study
  also examines the domain and mesh requirements necessary to capture
  the turbulent scales for the stable offshore boundary layers.
  Finally, we compare solutions computed using both Nalu-Wind and
  AMR-Wind solvers, and report on their relative accuracy and
  performance in this study.

@@

\section{Nomenclature}
\fxnote{Update symbols}
{\renewcommand\arraystretch{1.0}
  \noindent\begin{longtable*}{@{}l @{\quad=\quad} l@{}}
ABL       & Atmospheric Boundary Layer \\
$\alpha$  & Wind shear exponent  \\
$f$       & Frequency \\
$f_{max}$  & Maximum resolvable frequency due to mesh limits \\
$f_{Ny}$   & Nyquist frequency \\
$L$       & Turbulent integral length scale \\
$L_{Ob}$   & Obukhov length scale \\
LES       & Large Eddy Simulation \\
$u_\tau$   & Friction velocity    \\
$R_{ij}$   & Two point correlation tensor \\
SBL       & Stable Boundary Layer \\
$TI$      & Turbulence intensity \\
$TKE$     & Turbulent kinetic energy
\end{longtable*}}

@@

In addition to these measurement campaigns, large eddy simulations (LES) have
also been used to study ABL stability characteristics.  Recent work by
Kaul et al. \cite{kaul2020large} has shown that LES computations using
Nalu-Wind can capture the neutral and convectively unstable onshore
ABL.  Previous simulations by Cheung et al. \cite{cheung2020large}
successfully replicated the unstable and neutral ABL corresponding to
the Cape Wind meteorological tower measurements
\cite{archer2016predominance} and showed the effects of surface
heating on atmospheric stratification. LES of the stable ABL (SBL) presents
certain challenges compared to neutral and convective stability states, stemming
primarily from small dominant eddy length scales and low turbulence intensity.
These challenges require higher resolution grids and/or advanced subgrid
scale (SGS) models. As such, LES research of the SBL has been slower
to evolve than the other stability states.

The first successful LES of the SBL was published by Mason and Derbyshire in
1990 \cite{Mason1990}. They used a strategy of simulating a near steady-state
neutral stability turbulent simulation, then added a cooling-surface boundary
to develop the stable ABL. They recorded a need for implausibly strong initial
perturbations to sustain turbulence using the Smagorinsky SGS model. Brown et
al.\ \cite{Brown1994} worked with Mason and Derbyshire to add stochastic backscatter
to the SGS model, resulting in better agreement with surface layer observations.
Andren \cite{Andren1995}, simulated two SBL types, one with a neutral layer aloft, and
one capped by an inversion, comparing two SGS models. He found significant improvements
with the Sullivan SGS model, on par with the earlier backscatter model. Kemp and
Thompson \cite{Kemp1996} used the same model as \cite{Brown1994} to study particle
dispersion, but the LES code did not replicate an expected plume growth. Derbyshire
wrote a review of SBL modeling along with future projections in 1999 \cite{Derbyshire1999},
noting the importance of local Richardson number and identifying a need to go
to higher resolutions in SBL research.

In 2000, Kosovic and Curry \cite{Kosovic2000} used an LES simulation with a nonlinear SGS
model containing backscatter to favorably compare to the \fxnote{[Nate: define BASE]} BASE measurements, showing that
the SBL could reach a quasi steady-state. They also showed good agreement with the Nieustadt
analytical model. In the same year, Saiki et al. \cite{Saiki2000} simulated an SBL with
a strong capping inversion at z=500m, producing reasonable surface layer profiles and
identifying a need for SBL specific SGS models. Beare and MacVean \cite{Beare2004}
presented the first grid-focused work on LES of the SBL, using grids with deltas
between 2m and 12.5m, finding significant variation between grid sizes. The inclusion
of a backscatter model resulted in significant variation at a 6.25m resolution, but
very little at a 2m resolution. Basu et al. \cite{Basu2005} published a set of
turbulence statistics, measurements, and LES simulations in support
of the viability of Nieustadt's local scaling hypothesis under very stable
conditions. Jimenez and Cuxart \cite{Jimenez2005} set out to quantify the range
for which standard SGS models, using the Kolmogorov theory for dissipation, were
valid for increasingly stable flows. They found the model to perform adequately
for weakly stable, and some moderately stable flow regimes.

An inter-comparison of LES simulations of the SBL was organized in 2006, the
results of which are published in Beare et al.\ \cite{Beare2006}. Collectively,
this is known as the GABLS initiative, and involved simulating a moderately
stable flow taken from Kosovic and Curry \cite{Kosovic2000}. Eleven codes
were compared in all, wherein grid deltas of less than 12.5m were found to
successfully maintain resolved turbulence, grid deltas less than 3.125m
showed little statistical change, and results became independent of SGS
modeling at 1m resolution. In the same year, Kumar et al. \cite{Kumar2006}
successfully simulated an entire diurnal cycle with LES using a Lagrangian
dynamic scale dependent SGS model. This model has been used by several
researchers since 2011, including Lu and Porte-Agel \cite{Lu2011}, Huang
and Bou-Zeid \cite{Huang2013}, Park et al.\ \cite{Park2014}, and
Abkar and Porte-Agel \cite{Abkar2015}. Stoll and Porte-Agel \cite{Stoll2007}
compared three grid refinements and three scale dependent dynamic SGS models,
showing that the Lagrangian and locally averaged models reproduced important
flow features. The work of van Dop and Axelsen \cite{vanDop2007} provided
even more evidence in support Nieuwstadt's local scaling hypothesis for SBLs.
Chow and Zhou \cite{Chow2011} showed evidence that the Smagorinsky model is
simply too dissipative for use in strongly stable flows.

More recent work in LES of the SBL includes Matheou \cite{Matheou2016} who
explored a large parameter space using the Smagorinsky-Lilly SGS
closure, as well as Sullivan et al.\ \cite{sullivan2016turbulent}. Several researchers
have applied SBL simulation in the wind energy space \cite{Aitken2014, Abkar2015, Ghaisas2017}.
In 2020, Wurps et al. \cite{Wurps2020} provided a rigorous grid resolution study for the SBL,
looking at mean flow, turbulence variables, and two-point correlations, using a dynamic
SGS model that allows backscatter. They found grid requirements to be 2.5m, 10m, and 20m
for stable, neutral, and convective boundary layers respectively. They advocate looking
at two point correlations in order to quantify flow structure size, which in turn
provides a basis for evaluating grid resolution. In all of the reviewed literature, SBL LES
has been run in the onshore context only, a complete comparison
including offshore stable ABL conditions has yet to be completed.

\subsection{Objectives}
In the current study on stable marine boundary layers, we aim to build
on the previous work in the literature and complete several
objectives.  First, a series of stable boundary layer conditions will
be computed using the Exawind solvers which match the measurements of
Archer et al. \cite{archer2016predominance}.  The behavior of the
stable boundary layers, including mean flow quantities and turbulent
statistics, are examined and compared to the neutral and unstable
cases computed in \cite{cheung2020large}.  Secondly, we examine the
domain and mesh requirements necessary to capture the turbulent scales
for the stable offshore boundary layers.  Finally, two LES solvers
within the Exawind project are compared in this study: the
unstructured, incompressible Navier-Stokes solver Nalu-Wind and the
structured, adaptive mesh refinement solver AMR-Wind.  Both codes are
used to calculate the same SBL, and the relative differences between
the codes are reported.

### 2020-12-03 Paper/results.tex
Fix typos and stuff

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\caption{\label{tab:GridStudySetup} The setup for the grid resolution study} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & dx [m] & dy [m] & dz [m] & Domain size \\
  \hline
  Nalu-wind coarse  &  10.0  & 10.0   & 2.5 & 1.5km $\times$ 1.5km $\times$ 1.0km  \\
  Nalu-wind medium  &   5.0  &  5.0   & 2.5 & 1.5km $\times$ 1.5km $\times$ 1.0km  \\
  Nalu-wind fine    &   2.5  &  2.5   & 2.5 & 1.0km $\times$ 1.0km $\times$ 1.0km  \\
% AMR-wind fine     &   2.5  &  2.5   & 2.5 & 0.0 \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table} %[h!]
\caption{\label{tab:GridStudyLscale} The calculated turbulent integral
  lengthscale for each of the grid resolutions} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & dx [m] & Longitudinal L  & Lateral L \\
  \hline
  Nalu-wind coarse  &  10.0  & 36.4 m         & 0.0 m     \\
  Nalu-wind medium  &   5.0  & 21.5 m         & 4.52 m    \\
  Nalu-wind fine    &   2.5  & 25.3 m         & 5.92 m    \\
\hline
\end{tabular}
\end{table}

@@

%% %%%%%%%%%%% Compare Nalu/AMR lengthscale %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% % Postprocessing/ABLLength/CompareAMRNalu_ABL_Lengthscales.ipynb
%% \begin{figure} %[hbt!]
%%   \centering
%%   \fxnote{UPDATE THIS FIGURE}\\
%%   \includegraphics[width=2.5in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_Lengthscale_Stable_z20_Longitudinal.png}
%%   \includegraphics[width=2.5in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_Lengthscale_Stable_z20_Lateral.png}

@@

%%%%%%%%%%% z=20m snapshot pics %%%%%%%%%%%%%%%%%%%%%%%%%
% Created in Postprocessing/ABLStats/AMRWind_NaluWind_stable_AllWS_2p5Cubed.ipynb
\begin{figure}[hbt!]
  \centering
  %\fxnote{label plots using pstricks: 5m/s on top, 15m/s on bottom} \\
  \begin{tikzpicture}
    % 5 m/s plots
    \node[anchor=south west, inner sep=0](image) at (0in,4.50in)
         {\includegraphics[width=3in]{figures/snapshots/05ms/velh_mag_z20.png}};
    \node[align=left] at (1in, 4.50in){(a) horizontal vel, 5 m/s};
    \node[anchor=south west, inner sep=0](image) at (3in,4.35in)
         {\includegraphics[width=3.20in]{figures/snapshots/05ms/velz_z20_samelimits.png}};
    \node[align=left] at (4.25in, 4.50in){(b) vertical vel, 5 m/s};

>>>>

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\caption{\label{tab:GridStudySetup} The setup for the grid resolution study} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & dx [m] & dy [m] & dz [m] & Domain size \\
  \hline
  Nalu-Wind coarse  &  10.0  & 10.0   & 2.5 & 1.5km $\times$ 1.5km $\times$ 1.0km  \\
  Nalu-Wind medium  &   5.0  &  5.0   & 2.5 & 1.5km $\times$ 1.5km $\times$ 1.0km  \\
  Nalu-Wind fine    &   2.5  &  2.5   & 2.5 & 1.0km $\times$ 1.0km $\times$ 1.0km  \\
% AMR-wind fine     &   2.5  &  2.5   & 2.5 & 0.0 \\
\hline
\end{tabular}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%% GRID STUDY: INTEGRAL LENGTH %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table} %[h!]
\caption{\label{tab:GridStudyLscale} The calculated turbulent integral
  lengthscale for each of the grid resolutions} \centering
\begin{tabular}{ccccc}
  \hline
  Case              & dx [m] & Longitudinal L  & Lateral L \\
  \hline
  Nalu-Wind coarse  &  10.0  & 36.4 m         & 0.0 m     \\
  Nalu-Wind medium  &   5.0  & 21.5 m         & 4.52 m    \\
  Nalu-Wind fine    &   2.5  & 25.3 m         & 5.92 m    \\
\hline
\end{tabular}
\end{table}

@@

%% %%%%%%%%%%% Compare Nalu/AMR lengthscale %%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% % Postprocessing/ABLLength/CompareAMRNalu_ABL_Lengthscales.ipynb
%% \begin{figure} %[hbt!]
%%   \centering
%%   \includegraphics[width=2.5in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_Lengthscale_Stable_z20_Longitudinal.png}
%%   \includegraphics[width=2.5in]{figures/Compare_AMRWind_NaluWind/AMRWind_NaluWind_Lengthscale_Stable_z20_Lateral.png}

@@

%%%%%%%%%%% z=20m snapshot pics %%%%%%%%%%%%%%%%%%%%%%%%%
% Created in Postprocessing/ABLStats/AMRWind_NaluWind_stable_AllWS_2p5Cubed.ipynb
\begin{figure}[hbt!]
  \centering
  \begin{tikzpicture}
    % 5 m/s plots
    \node[anchor=south west, inner sep=0](image) at (0in,4.50in)
         {\includegraphics[width=3in]{figures/snapshots/05ms/velh_mag_z20.png}};
    \node[align=left] at (1in, 4.50in){(a) horizontal vel, 5 m/s};
    \node[anchor=south west, inner sep=0](image) at (3in,4.35in)
         {\includegraphics[width=3.20in]{figures/snapshots/05ms/velz_z20_samelimits.png}};
    \node[align=left] at (4.25in, 4.50in){(b) vertical vel, 5 m/s};
